Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering

### 2016-10-29 nips2016/review/rebuttal_xavier.txt

2763

We warmly thank the referees for their time and their helpful comments regarding our manuscript.

MNIST experiments (R1,R3,R7): The accuracy gap between standard CNN and graph CNN can be reduced with more epochs. With 20 epochs (instead of 10 in our submission) and LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%. This reveals a slower convergence of graph CNN, and future work will investigate better optimization strategies.

R1: Thanks for pointing out the GNN model (we will add this ref). GNN processes data in the spatial/graph domain, unlike our model that is defined in the spectral domain. It is not direct to formulate GNN in spectral form (and therefore to compare both techniques), that is, how to define spectral filters g s.t. Psi(x,y)=g(L)x (not only identity filter g(z)=z)? 
The spectral filters are indeed rotationally invariant. We think that this property might actually be an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN. We will explore this idea in future work.

R7: We indeed consider a single graph for this work, which is already meaningful for multiple applications s.a. brain connectivity analysis, and social network analysis (one graph and many data signals on this graph). Extension to multiple graphs is obviously of great interest, and can also be explored with a spectral technique like the one proposed here. We reserve this extension for another paper. 
The pooling strategy is inevitable in any CNN architecture as it allows to reduce spatial resolution and to increase filter resolution. In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering, which is NP-hard. We opted here for the Graclus model, which has proved to be a very efficient graph clustering technique.
It is expected that accuracy drops (3%) when the graph structure is destroyed with a random graph, as convolutional layers become meaningless. 
For 20NEWS, we compare against state-of-the-art Multinomial NB technique, and other baseline techniques.

R8: We have recently improved the presentation, accessibility, and Section 4 of the paper. We will be happy to release it.
When random graphs are used, then the convolutional layers become meaningless. However, the last part of the CNN system, the fully connected layers stay active, and absorbs the drop in performance. 
There are plenty of theoretical works in the area of spectral analysis, see e.g. the review [29].
The complexity of our technique is O(EK), which reduces to linear O(n) in case of sparse graphs, vs O(n^2) for [13].

R9: For the presentation of the paper, see our answer to R8.
We did not present any natural graph experiments due to the space limitation.

### 2016-10-29 nips2016/review/rebuttal.txt

6500

############
ALL
############

We warmly thank the referees for their time and their valuable comments regarding our manuscript.

We first address a common comment from Referees 1,3,7.

MNIST experiments: The accuracy gap between standard CNN and graph CNN can be reduced with more epochs. With 20 epochs (instead of 10 in our submission) and a LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%, bringing closer the accuracy gap between graph and classic CNN. This reveals a slower convergence of graph CNN, and future work will investigate better optimization strategies. ISOTROPIC?

We now address the specific comments of each referee.

Referee 1.
GNN: Thanks for pointing out the GNN model (we will add this ref). On the high level, GNN processes data in the spatial/graph domain, unlike our model that is defined in the spectral domain. Technically, GNN was originally designed to classify nodes or graphs by first embedding each node in an Euclidean space with RNNs, and then using those as features for classification or regression. Our technique is developed to classify graph signals by successively extracting higher level features, as done for classical CNN. Regarding the referee’s comment, we read very carefully the two 2009 GNN papers, and were not able to express the graph CNN in terms of GNN, as suggested by the referee. We would be grateful if the referee could point out more connections between our model and GNN.
Rotational invariance: The spectral filters are indeed rotationally invariant. Although it first appears to be a limitation of this system, we think the rational invariance property might actually be an advantage for large-scale image datasets. Our motivation is inspired by the following experiment (not reported in the paper): We learned the classic and graph CNN from a rotated MNIST dataset, that is each training MNIST image is randomly rotated from -90° to 90° (no data augmentation here, simply rotation of original images). Classic CNN obtained 70.65% of accuracy, and graph CNN got 91.47%. Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications where the training data are not enough to capture all meaningful features. We will investigate such idea in a future work, that is if such invariance property is an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN. 
We will also study the possibility to learn anisotropic Laplacians, and thus anisotropic filters.

Referee 3.
Table 4 XXX: The goal of this table was to show that convolutional layers, model CN32-softmax, are also effective on graphs. This is justified by comparing to fully connected networks, models FC. 
The performance of graph CNN did not overcome the state-of-the-art? text document classification technique, namely Multinomial NB. This may due to the small size of the dataset, 20NEWS only having 19K documents. CNN usually perform less than other classification techniques for small datasets. As a consequence, we will apply our technique to the larger text dataset RCV1 which has 800K documents, as [13].

Referee 7.
Single graph: We indeed consider a single graph in this work, which is already meaningful for multiple applications s.a. brain connectivity analysis, and social network analysis (one graph and many data signals on this graph). Transfer graph features/filters from one graph to another graph is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (for example learn and transfer Chebyshev coefficients across graphs). We reserve this extension for another paper. 
Graph pooling: Pooling is a pillar operation in any CNN architecture as it allows to reduce spatial resolution and increase feature resolution. In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering or coarsening, i.e. grouping similar vertices together, which is NP-hard. AGNOSTIC We opted here for the Metis/Graclus model [8,15], which has proved to be a very efficient graph clustering technique. Note also that the coarsening step is performed only once, before training.
Random graph: It is clearly expected that classification accuracy drops (3%) when the graph structure is inconsistent, i.e. for random graphs (graphs that hold random connexions between pixels), as convolutional layers become meaningless. It shows that the regular grid structure, exploiting by standard CNN, is obviously of first importance for images. Eventually, note that our technique has the flexibility to exploit additional information like pixel values through the arbitrary graph definition, unlike classic CNN.
20NEWS XXX: For this dataset, we compare against the state-of-the-art text document classification technique, namely Multinomial NB, and the SVM and linear classifier baseline techniques. As illustrated in Table 5, classification performances depend on the quality of the graph construction. However, we did not try to optimize the graph in our experiments. Pls, see also our answer to Referee 3.

Referee 8.
Presentation: We have recently improved the presentation and accessibility of the paper. We will be happy to release this new version in the final version of the paper.
Graph usefulness XXX: The goal of Section 6.2 and Table 5 is to show that the usefulness of the graph structure directly depends on the quality of the graph. When random graphs are used, then the convolutional layers become meaningless and graph CNN are not well suited for the problem at hand. On the contrary, if the graph is well-constructed, like the grid for MNIST, then graph CNN will provide good performances. 
Small performance drop: XXX For random graphs, convolutional layers are not useful. However, the last part of the CNN system, the fully connected layers stay active, and absorbs the drop in performance. 
Theory: XXX PERFORMANCES There are plenty of theoretical works in the area of spectral analysis for graphs and signal processing, see e.g. the review [29].
Comparison to [13]: Our technique is both faster and more accurate. See Table 6 with [13] being referred as Spline Filters. Besides, a theoretical advantage of our model is that our filters are exactly localized in K hops on graphs.

Referee 9.
Presentation: See our answer to Referee 8.
Natural graphs: Indeed, we did not present any natural graph experiments in this work. Our first goal was to introduce our efficient graph CNN technique. However, a data analysis project on Wikipedia using this technique is ongoing.

############
Private to AC
############

The general consensus of the referees lean towards acceptance. Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful. The only concerns come from Referees 8,9 whom acknowledge their low confidence in the understanding of the main contributions of the paper. XXX Besides, questions of Referee 8 about theoretical background and effectiveness of the algorithm are repeatedly provided throughout the paper like in Section 2 and Section 6.3.

### 2016-10-29 nips2016/review/rebuttal.txt

The general consensus of the referees lean towards acceptance. Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful. The only concerns come from Referees 8,9 who acknowledge their low confidence in the understanding of the main contributions of the paper.

@@

############
ALL
############

5900

We warmly thank the referees for their time and their valuable comments regarding our manuscript.

We first address common comments.

Accuracy Gap on MNIST (Referees 1,3,7): There is indeed an accuracy gap between standard CNN and graph CNN. It may due to several reasons such as architecture tuning and isotopic filters. We also observed that the accuracy gap can be reduced with more epochs. With 20 epochs (instead of 10 in our submission) and a LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%. This reveals a slower convergence of graph CNN, and we will investigate better optimization strategies.

Graph Usefulness and Experiments (R3,7,8): The goal of Tables 3 and 5 is to show that the usefulness of the graph structure directly depends on the quality of the graph. When random graphs are used (graphs that hold random connexions between pixels/vertices), then the convolutional layers become meaningless, and graph CNN are not well suited for the problem at hand. On the contrary, if the graph is well constructed, like the regular grid for MNIST, then graph CNN provide good performances. For 20NEWS, we did not try to optimize the graph structure in our experiments. There may be better ways to construct higher-quality graphs. However, constructing graphs from data is an open and challenging question, and was not our objective here. 
The goal of Table 4 was to demonstrate that convolutional layers, model CN32-softmax, are also effective on graphs that are not regular image grids, but a graph of 20NEWS words. This is justified by comparing CN32-softmax to fully connected networks, FC models. We also compared against the effective and popular text document classification technique, Multinomial NB, and the SVM and linear classifier baseline techniques. Graph CNN did not overcome Multinomial NB. It may due to the small size of the dataset, 20NEWS having only 19K documents. CNN usually perform less than other classification techniques for small datasets. As a consequence, we will apply our technique to the larger text dataset RCV1 which has 800K documents, as [13].

@@

We now address the specific comments of each referee.

Referee 1.
GNN: Thanks for pointing out the GNN model (we will add this ref). On the high level, GNN is a spatial/graph technique, unlike our model which is spectral. Technically, GNN was originally designed to classify nodes or graphs by first embedding each node in an Euclidean space with RNN, and then using those as features for classification or regression. Our technique is developed to classify graph signals by successively extracting higher level features, as done for classical CNN. Regarding the referee’s comment, we read very carefully the two 2009 GNN papers, and were not able to express the graph CNN in terms of GNN, as suggested by the referee. We would be grateful if the referee could point out more connections between our model and the GNN papers.
Rotational invariance: The spectral filters are indeed rotationally invariant. Although it first appears to be a limitation of this system, we think the rational invariance property might actually be an advantage for large-scale image datasets. Our motivation is inspired by the following experiment (not reported in the paper). We learned the classic and graph CNN from a rotated MNIST dataset, that is each training MNIST image is randomly rotated from -90° to 90° (no data augmentation here, simply rotation of original images). Classic CNN obtained 70.65% of accuracy, and graph CNN got 91.47%. Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications where the training data are not enough to capture all meaningful features. We will investigate such idea in a future work, that is if such invariance property is an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN. 
We will also study the possibility to learn anisotropic Laplacians, and thus anisotropic filters.

Referee 7.
Single graph: We indeed consider a single graph in this work, which is already meaningful for multiple applications s.a. brain connectivity analysis, and social network analysis (one graph and many data signals on this graph). Transfer graph features/filters from one graph to another graph is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (for example learn and transfer Chebyshev coefficients across graphs). We reserve this extension for another paper. 
Graph pooling: Pooling is a pillar operation in any CNN architecture as it allows to reduce spatial resolution and increase feature resolution. In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering or coarsening, i.e. grouping similar vertices together, which is NP-hard. We opted here for the Metis/Graclus model [8,15], which has proved to be a very efficient graph clustering technique. Note also that the coarsening step is performed only once, before training.

Referee 8.
Presentation: We have recently improved the presentation and accessibility of the paper. We will be happy to release this new version in the final version of the paper.
Theory: The analysis of the performances of standard CNN is currently a central and open mathematical question. It is thus challenging to state anything about the theoretical performances of our model vs [13]. The only performance comparison is experimental and are summarised in Table 6. Besides, a theoretical advantage of our model is that our filters are exactly localized in K hops on graphs.

Referee 9.
Presentation: See our answer to R8.
Natural graphs: Indeed, we did not present any natural graph experiments in this work. Our first goal was to introduce our efficient graph CNN technique. However, a data analysis project on Wikipedia network using this technique is ongoing.

### 2016-10-29 nips2016/review/rebuttal.txt

4950

We warmly thank the referees for their time and valuable comments.

Referees 1,3,7
Accuracy Gap for MNIST: The gap between standard CNN and graph CNN may be due to our isotopic filters. We also observed that with 20 epochs (instead of 10 in our submission) for LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%. This reveals a slower convergence of graph CNN, and the need to investigate better optimization strategies.

Referees 3,7,8
Graph Usefulness and Experiments: The goal of Tables 3,5 is to show that the usefulness of the graph structure directly depends on the graph quality. When random graphs are used, then convolutional layers are useless, and graph CNN are not well suited for the problem. If the graph is well constructed, like the grid for MNIST, then graph CNN provide good performances. For 20NEWS, we did not try to optimize the graph structure. There may be better ways to construct higher-quality graphs, but it was not our objective. 
The goal of Table 4 is to demonstrate that convolutional layers CN32 are also effective on graphs that are not regular image grids. This is showed by comparing CN32 to fully connected (FC) networks. We also compared against the effective Multinomial NB document classification technique, and SVM and linear classifier baseline techniques. Graph CNN did not overcome Multinomial NB. It may be due to the small size of the dataset (20NEWS has 19K documents). CNN usually perform less than other classification techniques for small datasets. We plan to apply our technique to the larger text dataset RCV1 (800K documents).

Referee 1
Thanks for pointing out GNN (we will add this ref). On the high level, GNN is a spatial technique, and our model is spectral. GNN was designed to classify nodes/graphs by first embedding each node in an Euclidean space with RNN, and then using those as features for classification. Our technique is developed to classify graph signals by successively extracting higher level features, as done for classical CNN. Regarding the referee’s comment, we read carefully the two GNN papers, and were not able to express the graph CNN in terms of GNN, as suggested by the referee. We would be grateful if the referee could point out more connections between our model and the GNN papers.
The spectral filters are indeed rotationally invariant. Although it first appears to be a limitation, we think the rotational invariance property might actually be an advantage. Our motivation is inspired by the following experiment (not reported in the paper). We learned the classic and graph CNN from a rotated MNIST dataset, i.e. each training image is randomly rotated (no data augmentation here, simply rotation of original images). Classic CNN obtained 70% of accuracy, and graph CNN 91%. Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications where the training data are not enough to capture all meaningful features. We will investigate such idea in a future work, that is if such invariance property is an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN. 
We will also study the possibility to learn anisotropic Laplacians, and thus anisotropic filters.

Referee 7
We indeed consider a single graph in this work, which is already meaningful for multiple applications s.a. brain connectivity analysis, and social network analysis. Transfer graph features/filters from one graph to another graph is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (e.g. learn and transfer Chebyshev coefficients across graphs). We reserve this extension for another paper. 
Pooling is a pillar operation in any CNN architecture as it allows to reduce spatial resolution and increase feature resolution. In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering or coarsening, i.e. grouping similar vertices together, which is NP-hard. We opted here for the Metis model, which has proved to be very efficient. Note also that the coarsening step is performed only once.

Referee 8
We have recently improved the presentation of the paper, and will release the new version in the final version of the paper.
The performance analysis of standard CNN is a central and open mathematical question. It is thus challenging to state anything about the theoretical performances of our model. The only performance study is experimental and summarised in Table 6.

Referee 9
Presentation: See our answer to R8.
We indeed did not introduce any natural graph experiments in this work. Our first goal was to introduce an efficient graph CNN technique. However, a data analysis project on Wikipedia network using this technique is ongoing.

@@

Referee 9
Presentation: See our answer to R8.
We indeed did not introduce any natural graph experiments in this work. Our first goal was to introduce an efficient graph CNN technique. However, a data analysis project on Wikipedia network using this technique is ongoing.

############
Private to AC
############

The general consensus of the referees lean towards acceptance. Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful. The only concerns come from Referees 8,9 who acknowledge their low confidence in their understanding of the paper main contributions.

### 2016-10-29 nips2016/review/rebuttal.txt

############
ALL
############

4992

We warmly thank the referees for their time and valuable comments.

Referees 1,3,7
While the isotropic nature of our filters (see answer to R1) and the lack of experience on architecture design are certainly explanations for the gap between classical CNNs and graph CNNs on MNIST, we observed that with 20 epochs (instead of 10 in our submission) for the same LeNet5 architectures, classical CNNs obtained 99.33% accuracy and graph CNNs 99.14%. This reveals a slower convergence of graph CNNs, and the need to investigate better suited optimization or initialization strategies.

Referees 3,7,8
The goal of Tables 3 and 5 is to show that the usefulness of the graph structure directly depends on the graph quality. Random graphs are used as a baseline to assess graph quality, as convolutional layers are almost useless in this setting. A well suited graph, like the grid for MNIST, is useful and allows our model to perform well. We did however not try to optimize the graph structure for 20NEWS: there may be better ways to construct higher-quality graphs. It was not our objective, which was to show how graph quality impacts performance.
The goal of Table 4 is to demonstrate that a convolutional layer is also effective on graphs that are not regular grids, by comparing it to fully connected networks. We also compared against the effective Multinomial Naive Bayes and two baselines. Graph CNNs did not overcome Multinomial NB, likely because of the small size of the 20NEWS dataset (~19k documents). We plan to apply our technique to the larger RCV1 dataset (800k documents).

Referee 1
Thanks for pointing out GNN (we will add this ref). At a high level, GNN was designed to classify nodes or graphs by first embedding each node in an Euclidean space with an RNN, then using those as features for classification/regression. Our technique was developed to classify graph signals by successively extracting higher level features, similarly to classical CNNs. After reading the paper carefully, we were not able to express our graph CNN in terms of a GNN, as suggested by the referee. We would be grateful if the referee could point out more connections between our model and the GNN papers.
The spectral filters are indeed rotationally invariant. Whether this is a limitation or an advantage depends on the problem and should be verified empirically, as for any other invariance. Our claim is motivated by the following experiment (not reported in the paper): on a rotated MNIST dataset, i.e. each image is randomly rotated (no data augmentation here), our model obtained an accuracy of 91% while classical CNNs obtained 70%. Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications. Moreover, rotational invariance has been sought in the past. Many data augmentation schemes have used rotated versions of (part of) images to force a network to learn this invariance. Spatial Transformer Networks have been developed to learn invariance to translation, scale, rotation and more generic warping. In a future work, we will test our model on large-scale image datasets and investigate if such an invariance is an advantage. We will also study the possibility to learn anisotropic filters with the help of anisotropic Laplacians.

Referee 7
We indeed consider a single graph in this work, which is already meaningful for multiple applications such as brain connectivity analysis, documents network analysis and social network analysis. Transferring graph features/filters from one graph to another is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (e.g. learn and transfer Chebyshev coefficients across graphs). We reserve this extension for another paper.
A pooling strategy will always require some sort of coarsening, i.e. a way to group nodes to form a receptive field. While grouping is straightforward on grids, clustering or coarsening on arbitrary graphs is NP-hard. We opted here for the greedy Graclus algorithm, which has proved to be very efficient. Note that our method is however agnostic to the chosen coarsening algorithm and that coarsening is performed only once.

Referee 8
We did recently improve the presentation of the paper. Expect the final version .
The performance analysis of classical CNNs is a central and open mathematical question. It is thus challenging to state anything about the theoretical performances of our model. The only performance study is experimental and summarized in Table 6, which shows that our model is indeed superior to [13].

Referee 9
Presentation: see our answer to R8.
We did not introduce any natural graph experiments in this work. Our first goal was to introduce an efficient extension of CNNs to graphs and demonstrate its usefulness. Constructed graphs are a great way of demonstrating this. However, a data analysis project on the network of Wikipedia articles using this technique is ongoing.

@@

############
Private to AC
############

The general consensus of the referees lean towards acceptance. Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful. The only concerns come from Referees 8,9 who acknowledge their low confidence in their understanding of the paper main contributions.

Please note that, to foster innovation, we are going to publish the code used to produce our results once the paper is published.
